Building Sustainable Software - The CSDMS Approach

### 2014-07-14 hutton_et_al_2014_wssspe.tex
Added Albert's changes.

\author{
  Eric W. H. Hutton$^{\dag}$,
  Mark D. Piper$^{\dag}$,
  Irina Overeem$^{\dag}$,
  Albert J. Kettner$^{\dag}$
}

@@

\csdms{}, The Community Surface Dynamics Modeling System, is an NSF funded
project whose focus is to aid a diverse community of earth and ocean system
model \emph{users} and \emph{developers} to use and create robust software
quickly.  To this end, \csdms{} develops, integrates, archives and disseminates
earth-system models and tools to an international (67 country) community
with the goal of building the set of tools necessary to model the
earth system. Modelers look to \csdms{} for access to hundreds of open
source surface-dynamics models and tools, as well as model metadata. Such a
model repository increases model transparency and helps eliminate
duplication by presenting the current state of modeling efforts.
To increase software sustainability, composability and interoperability,
\csdms{} promotes standards that define common modeling interfaces, semantic
mediation between models, and model metadata. Through online resources and
workshops, \csdms{} promotes software engineering best practices, which are
unfamiliar to many developers within our modeling community. For example,
version control, unit testing, continious integration, test-driven
development, and well-written \emph{clean code} are all topics of the
educational mission of \csdms{}.

@@

\csdms{} has implemented a
Digital Object Identifier (\doi{}) system for software within the model
repository to guarantee recognition to software contributors.
\csdms{} was among the first software venues to assign \doi{}s to
open source software. The advantages of adopting a \doi{} system for open
source software are: 
\begin{itemize}
\item Guarantee credit to the software developer
\item Easily reuse and replicate research as software is directly locatable 
\item Increase software visibility: \doi{} content is 5 times more likely to
      deliver active links than content without.
\item Provide funding agencies with the ability to track usage, to measure
      impact.
\end{itemize}

@@

The \bmi{} identifies the entry points into software components that provide a
calling application with the necessary level of control to connect components.
\csdms{}, as well as other modeling frameworks such as
ESMF~\cite{hill2004architecture}, OpenMI~\cite{gregersen2007openmi}, and
OMS~\cite{david2002object}, identifies an interface that, at a minimum,
provides functionality to \emph{initialize}, \emph{update}, and
\emph{finalize} a component. However, for more complicated (and 2-way)
coupling components must implement the \bmi{} methods providing data access,
and model metadata.

@@

\csdms{} constantly evaluates the effectiveness of its tools and standards
through user feedback from clinics and tutorials and one-on-one interviews
with \csdms{} members.  Modeling courses at the University of Colorado,
clinics at the International Summer Institute at the University of Minnesota,
and clinics at annual \csdms{} Annual Meetings are the dominant forums for
this testing and feedback.

Over the last 4 years over 150 graduate students - from geography, hydrology,
geology, oceanography, earth sciences and engineering departments - have
enrolled in intensive \csdms{}-run modeling courses. These courses
provide an excellent use case for us to test usability and access to our
tools. Students reported in post-course surveys that they were unfamiliar with
the use of models on a high performance computing cluster beforehand (rated
their familiarity as 1.5 average on a scale of 1-5), but ended the courses
with familiarity and being more comfortable with running and changing
simulations as well as visualizing results (average rating approaches 4 on
a Likert scale of 1-5)~\cite{overeem2013strategies}.

>>>>

\author{
  Eric W. H. Hutton$^{\dag}$,
  Mark D. Piper$^{\dag}$,
  Irina Overeem$^{\dag}$,
  Albert J. Kettner$^{\dag}$,
  James P. M. Syvitski$^{\dag}$,
}

@@

\csdms{}, The Community Surface Dynamics Modeling System, is an NSF funded
project whose focus is to aid a diverse community of earth and ocean system
model \emph{users} and \emph{developers} to use and create robust software
quickly.  To this end, \csdms{} develops, integrates, archives and disseminates
earth-system models and tools to an international (67 country) community
with the goal of building the set of tools necessary to model the
earth system. Modelers use \csdms{} for access to hundreds of open
source surface-dynamics models and tools, as well as model metadata. Such a
model repository increases model transparency and helps eliminate
duplication by presenting the current state of modeling efforts.
To increase software sustainability, composability and interoperability,
\csdms{} promotes standards that define common modeling interfaces, semantic
mediation between models, and model metadata. Through online resources and
workshops, \csdms{} promotes software engineering best practices, which are
unfamiliar to many developers within our modeling community. For example,
version control, unit testing, continious integration, test-driven
development, and well-written \emph{clean code} are all topics of the
educational mission of \csdms{}.

@@

\csdms{} has implemented a
Digital Object Identifier (\doi{}) system for software within its model
repository to guarantee recognition to software contributors.
\csdms{} was among the first software venues to assign \doi{}s to
open source software. The advantages of adopting a \doi{} system for open
source software are: 
\begin{itemize}
\item Guarantee credit to the software developer
\item Easily reuse and replicate research as software is directly locatable 
\item Increase software visibility: \doi{} content is 5 times more likely to
      deliver active links than content without.
\item Provide funding agencies with the ability to track usage, to measure
      impact.
\end{itemize}

@@

The \bmi{} identifies the entry points into software components that provide a
calling application with the necessary level of control to connect components.
\csdms{}, as well as other modeling frameworks such as
ESMF~\cite{hill2004architecture}, OpenMI~\cite{gregersen2007openmi}, and
OMS~\cite{david2002object}, identifies an interface that, at a minimum,
provides functionality to \emph{initialize}, \emph{update}, and
\emph{finalize} a component. For more complicated (and 2-way)
coupling, components must implement \bmi{} methods providing data access,
and model metadata. For data access, \bmi{} defines a set of \emph{getter}
and \emph{setter} methods that allow model components to present internal data
for both reading and writing.

@@

\csdms{} constantly evaluates the effectiveness of its tools and standards
through user feedback from clinics and tutorials and one-on-one interviews
with \csdms{} members.  Modeling courses at the University of Colorado,
clinics at the International Summer Institute at the University of Minnesota,
and clinics at \csdms{} Annual Meetings are the dominant forums for
this testing and feedback.

Over the last 4 years over 150 graduate students - from geography, hydrology,
geology, oceanography, earth sciences and engineering departments - have
enrolled in intensive \csdms{}-run modeling courses. These courses
provide an excellent use case for us to test usability and access to our
tools. Students reported in post-course surveys that they were unfamiliar with
the use of models on a high performance computing cluster beforehand (rated
their familiarity as 1.5 average on a Likert scale of 1-5), but ended the
courses with familiarity and being more comfortable with running and changing
simulations as well as visualizing results (average rating approaches 4)
~\cite{overeem2013strategies}.
